Uncertainty propagation in neural networks for sparse coding

### 2018-11-27 nips_workshop_2018/poster/BayesianLISTAposter.tex
upd poster
upd presentation

@@

        \begin{algorithm}[H]
          \begin{algorithmic}
            \STATE \textit{Init.} Dense $\mathbf{b} \gets \mathbf{W}\mathbf{y}$
            \STATE \textit{Init.} Soft-thresholding $\widehat{\boldsymbol\beta}_0 \gets h_\lambda(\mathbf{b})$
            \FOR{$l=1$ \TO $L$} 
            \STATE Dense $\mathbf{c}_l \gets \mathbf{b} + \mathbf{S}\widehat{\boldsymbol\beta}_{l-1}$
            \STATE Soft-thresholding $\widehat{\boldsymbol\beta}_{l} \gets h_\lambda(\mathbf{c}_l)$
            \ENDFOR
            \RETURN $\widehat{\boldsymbol\beta} \gets \widehat{\boldsymbol\beta}_{L}$
          \end{algorithmic}   
        \end{algorithm}
          \begin{itemize}
            \item[] {\color{red}Overfitting}
            \item[] {\color{red}No uncertainty estimation}
          \end{itemize} 
      \end{minipage} 
      \begin{minipage}[t]{0.4\linewidth} 
        \begin{tikzfigure}
          \includegraphics[width=400pt]{graphics/LISTA_main.pdf}
        \end{tikzfigure}
      \end{minipage}
    }
    \block{3-Uncertainty propagation}{
      \innerblock{Idea}{
        At every step the output of soft-thresholding can be closely approximated with spike and slab distribution
      \begin{enumerate}
        \item \(\mathbf{b} = \mathbf{W}\mathbf{y}\) is Gaussian-distributed
        \item \(\widehat{\boldsymbol\beta}_{0} = h_\lambda(\mathbf{b})\) is approximated with the spike and slab distribution
        \item \(\mathbf{e}_l = \mathbf{S}\widehat{\boldsymbol\beta}_{l-1}\) is approximated with the Gaussian distribution
        \item \(\mathbf{c}_l = \mathbf{b} + \mathbf{e}_l\) is Gaussian-distributed
        \item \(\widehat{\boldsymbol\beta}_{l} = h_\lambda(\mathbf{c}_l)\) is approximated with the spike and slab distribution
      \end{enumerate}
      }
      \innerblock{Advantages}{
        \begin{itemize}
          \item All latent variables are modelled with parametrised distributions
          \item We can apply approximate Bayesian inference methods
        \end{itemize}
      }
    }
    \column{0.5}
    \block{2-BayesLISTA}{ 
      \innerblock{Bayesian LISTA}{
      \begin{itemize}
        \item[] Add priors for NN weights
        \begin{equation}
          p(\mathbf{W}) = \prod_{d=1}^D\prod_{k=1}^K \mathcal{N}(w_{ij} ; 0, \eta^{-1}), \quad
          p(\mathbf{S}) = \prod_{d'=1}^D\prod_{d''=1}^D \mathcal{N}(s_{d'd''} ; 0, \eta^{-1}),
        \end{equation}
        \item[] Propagate distribution for $\widehat{\boldsymbol\beta}$ through layers
        \item[] Compute prediction as noisy NN output
        \begin{equation}
          p(\mathbf{\boldsymbol\beta}| \mathbf{y}, \mathbf{W}, \mathbf{S}, \gamma, \lambda)
          = \prod_{d=1}^D\mathcal{N}\left(\beta_d; [f(\mathbf{y} ; \mathbf{S}, \mathbf{W}, \lambda)]_d, \gamma^{-1}\right)
        \end{equation}
        \item[] Update weights with PBP
      \end{itemize}
      }
    }
    \block{4-BackProp-PBP}{Blocktext}
    \block{5-Results}{
      \innerblock{Synthetic Experiments}{
        \begin{tikzfigure}[Different depth performance]
          \subcaptionbox*{\emph{NMSE}}{%
                     \includegraphics[width=200pt]{graphics/synthetic_number_of_layers/nmse_validation}
          }
          \quad
          \subcaptionbox*{\emph{F measure}}{%
                     \includegraphics[width=200pt]{graphics/synthetic_number_of_layers/f_measure_validation}
          }
          %\captionsetup{labelformat=empty}
        \end{tikzfigure}

        \begin{tikzfigure}[Different observation size performance]
          \subcaptionbox*{\emph{NMSE}}{%
                     \includegraphics[width=200pt]{graphics/synthetic_undersampling/nmse_validation}
          }
          \quad
          \subcaptionbox*{\emph{F measure}}{%
                     \includegraphics[width=200pt]{graphics/synthetic_undersampling/f_measure_validation}
          }
          %\captionsetup{labelformat=empty}
        \end{tikzfigure}
      }

      \innerblock{MNIST Experiments}{
        \begin{tikzfigure}[Posterior parameters for an image of digit 7]
          \subcaptionbox*{posterior mean}{
            \includegraphics[width=200pt]{graphics/posterior_mean}
          }
          \quad
          \subcaptionbox*{posterior spike indicator}{
            \includegraphics[width=200pt]{graphics/posterior_spike_indicator}
          }
        \end{tikzfigure}

        \begin{tikzfigure}[Samples from the posterior for an image of digit 7]
          \subcaptionbox*{posterior sample 0}{
            \includegraphics[width=150pt]{graphics/posterior_sample_0}
          }
          \quad
          \subcaptionbox*{posterior sample 1}{
            \includegraphics[width=150pt]{graphics/posterior_sample_1}
          }
          \quad
          \subcaptionbox*{posterior sample 2}{
            \includegraphics[width=150pt]{graphics/posterior_sample_2}
          }
        \end{tikzfigure}
      }

      \innerblock{Active Learning}{
        Use the estimated uncertainty to choose next training data with largest variance
        \begin{tikzfigure}[Sequential pool additions]
        \subcaptionbox*{NMSE}{
            \includegraphics[width=200pt]{graphics/active_mnist/nmse_validation}
        }
        \subcaptionbox*{F measure}{
            \includegraphics[width=200pt]{graphics/active_mnist/f_measure_validation}
        }
        \end{tikzfigure}
      }
    }

    \block{6-Summary}{Blocktext}
    \note{Notetext}
  \end{columns}
\end{document}

>>>>

\renewenvironment{tikzfigure}[1][]{
  \def \rememberparameter{#1}
  \vspace{10pt}
  \refstepcounter{figurecounter}
  \begin{center}
  }{
    \ifx\rememberparameter\@empty
    \else %nothing
    \\[10pt]
    %{\small Fig.~\thefigurecounter: \rememberparameter}
    {\small \rememberparameter}
    \fi
  \end{center}
}

@@

        \begin{algorithm}[H]
          \begin{algorithmic}
            \STATE \textit{Init.} Dense $\mathbf{b} \gets \mathbf{W}\mathbf{y}$
            \STATE \textit{Init.} Soft-thresholding $\widehat{\boldsymbol\beta}_0 \gets h_\lambda(\mathbf{b})$
            \FOR{$l=1$ \TO $L$} 
            \STATE Dense $\mathbf{c}_l \gets \mathbf{b} + \mathbf{S}\widehat{\boldsymbol\beta}_{l-1}$
            \STATE Soft-thresholding $\widehat{\boldsymbol\beta}_{l} \gets h_\lambda(\mathbf{c}_l)$
            \ENDFOR
            \RETURN $\widehat{\boldsymbol\beta} \gets \widehat{\boldsymbol\beta}_{L}$
          \end{algorithmic}   
        \end{algorithm}
          \begin{itemize}
            \item[] {\color{red}Overfitting}
            \item[] {\color{red}No uncertainty estimation}
          \end{itemize} 
      \end{minipage} 
      \begin{minipage}[t]{0.4\linewidth} 
        \begin{tikzfigure}
          \includegraphics[width=400pt]{graphics/LISTA_main.pdf}
        \end{tikzfigure}
      \end{minipage}
    }
    \block{3-Uncertainty propagation}{
      %\innerblock{Idea}{
        At every step the output of soft-thresholding can be closely approximated with spike and slab distribution
      \begin{enumerate}
        \item \(\mathbf{b} = \mathbf{W}\mathbf{y}\) is Gaussian-distributed
        \item \(\widehat{\boldsymbol\beta}_{0} = h_\lambda(\mathbf{b})\) is approximated with the spike and slab distribution
        \item \(\mathbf{e}_l = \mathbf{S}\widehat{\boldsymbol\beta}_{l-1}\) is approximated with the Gaussian distribution
        \item \(\mathbf{c}_l = \mathbf{b} + \mathbf{e}_l\) is Gaussian-distributed
        \item \(\widehat{\boldsymbol\beta}_{l} = h_\lambda(\mathbf{c}_l)\) is approximated with the spike and slab distribution
      \end{enumerate}
      \begin{tikzfigure}
        \includegraphics[width=400pt]{graphics/spsl_propagation.pdf}
      \end{tikzfigure}
      %}
      %\innerblock{Advantages}{
        \begin{itemize}
          \item All latent variables are modelled with parametrised distributions
          \item We can apply approximate Bayesian inference methods
        \end{itemize}
      %}
    }
    \block{6-References}{
      \begin{itemize}
        \item J. M. Hern{\'a}ndez-Lobato, R. Adams. Probabilistic backpropagation for scalable learning of Bayesian neural networks. ICML 2015.
        \item K. Gregor, Y. LeCun. Learning fast approximations of sparse coding. ICML 2010.
      \end{itemize}
    }
    \column{0.5}
    \block{2-BayesLISTA}{ 
      \begin{itemize}
        \item Add priors for NN weights
        \begin{equation}
          p(\mathbf{W}) = \prod_{d=1}^D\prod_{k=1}^K \mathcal{N}(w_{ij} ; 0, \eta^{-1}), \quad
          p(\mathbf{S}) = \prod_{d'=1}^D\prod_{d''=1}^D \mathcal{N}(s_{d'd''} ; 0, \eta^{-1}),
        \end{equation}
        \item Propagate distribution for $\widehat{\boldsymbol\beta}$ through layers
        \item Compute prediction as noisy NN output
        \begin{equation}
          p(\mathbf{\boldsymbol\beta}| \mathbf{y}, \mathbf{W}, \mathbf{S}, \gamma, \lambda)
          = \prod_{d=1}^D\mathcal{N}\left(\beta_d; [f(\mathbf{y} ; \mathbf{S}, \mathbf{W}, \lambda)]_d, \gamma^{-1}\right)
        \end{equation}
        \item Update weights with PBP
      \end{itemize}
      }
    \block{4-BackProp-PBP}{
      Approximate posterior
    \begin{align}
    \label{eq:approximating_dsitribution}
    \begin{split}
    q(\mathbf{W}, \mathbf{S}, \gamma, \eta) &= \prod_{d=1}^D\prod_{k=1}^K \mathcal{N}(w_{dk} ; m^w_{dk}, v^w_{dk}) \prod_{d'=1}^D\prod_{d''=1}^D \mathcal{N}(s_{d'd''} ; m^s_{d'd''}, v^s_{d'd''}) \\
    &\times \text{Gam}(\gamma; a^\gamma, b^\gamma) \text{Gam}(\eta; a^\eta, b^\eta)
    \end{split}
    \end{align}

    probabilistic backpropagation algorithm: derivatives of the logarithm of a normalisation constant

    \begin{equation}
    q(a) = Z^{-1}f(a)\mathcal{N}(a; m, v)
    \end{equation}
    \begin{equation}
    Z \approx \prod_{d=1}^D \left[\omega^{\widehat{\boldsymbol\beta}}_d  \mathcal{T}\left(\beta_d ; 0, \beta^\gamma / \alpha^\gamma, 2\alpha^\gamma\right) + \vphantom{m^{\widehat{\boldsymbol\beta}}_d} \left(1 - \omega^{\widehat{\boldsymbol\beta}}_d\right)\mathcal{N}\left(\beta_d ; m^{\widehat{\boldsymbol\beta}}_d,  \beta^\gamma / (\alpha^\gamma - 1) + v^{\widehat{\boldsymbol\beta}}_d\right)\right],
    \end{equation}
    where $\{\omega^{\widehat{\boldsymbol\beta}}_d, m^{\widehat{\boldsymbol\beta}}_d, v^{\widehat{\boldsymbol\beta}}_d\}$ are the parameters of the spike and slab distribution for $[\widehat{\boldsymbol\beta}]_d$.

    }
    \block{5-Results}{
      %\innerblock{Synthetic Experiments}{
        \emph{Synthetic}
        \begin{tikzfigure}[Different depth performance]
          \subcaptionbox*{\emph{NMSE}}{%
                     \includegraphics[width=300pt]{graphics/synthetic_number_of_layers/nmse_validation}
          }
          \quad
          \subcaptionbox*{\emph{F measure}}{%
                     \includegraphics[width=300pt]{graphics/synthetic_number_of_layers/f_measure_validation}
          }
          %\captionsetup{labelformat=empty}
        \end{tikzfigure}

        \begin{tikzfigure}[Different observation size performance]
          \subcaptionbox*{\emph{NMSE}}{%
                     \includegraphics[width=300pt]{graphics/synthetic_undersampling/nmse_validation}
          }
          \quad
          \subcaptionbox*{\emph{F measure}}{%
                     \includegraphics[width=300pt]{graphics/synthetic_undersampling/f_measure_validation}
          }
          %\captionsetup{labelformat=empty}
        \end{tikzfigure}
      %}
      \emph{MNIST}
      %\innerblock{MNIST Experiments}{
        \begin{tikzfigure}[Posterior parameters for an image of digit 7]
          \subcaptionbox*{posterior mean}{
            \includegraphics[width=200pt]{graphics/posterior_mean}
          }
          \quad
          \subcaptionbox*{posterior spike indicator}{
            \includegraphics[width=200pt]{graphics/posterior_spike_indicator}
          }
        \end{tikzfigure}

        \begin{tikzfigure}[Samples from the posterior for an image of digit 7]
          \subcaptionbox*{posterior sample 0}{
            \includegraphics[width=200pt]{graphics/posterior_sample_0}
          }
          \quad
          \subcaptionbox*{posterior sample 1}{
            \includegraphics[width=200pt]{graphics/posterior_sample_1}
          }
          \quad
          \subcaptionbox*{posterior sample 2}{
            \includegraphics[width=200pt]{graphics/posterior_sample_2}
          }
        \end{tikzfigure}
      %}

      %\innerblock{Active Learning}{
      \emph{Active Learning}
        Use the estimated uncertainty to choose next training data with largest variance
        \begin{tikzfigure}[Sequential pool additions]
        \subcaptionbox*{NMSE}{
            \includegraphics[width=300pt]{graphics/active_mnist/nmse_validation}
        }
        \subcaptionbox*{F measure}{
            \includegraphics[width=300pt]{graphics/active_mnist/f_measure_validation}
        }
        \end{tikzfigure}
      %}

    }
  \end{columns}
\end{document}
